Solving Collective Spin Hamiltonians with Decoherence

### 2018-09-07 soc_squeezing/cumulant_expansions.tex
added alternate method for computing arbitrary single-spin decoherence

The most general type of single-spin decoherence is described by
\begin{align}
  \D\p{\gamma;\O}
  = \sum_j\p{{\gamma^{(j)}}^\dag \O \gamma^{(j)}
    - \f12\sp{{\gamma^{(j)}}^\dag \gamma^{(j)}, \O}_+},
  &&
  \gamma \equiv \gamma_\z \sigma_\z
  + \gamma_+ \sigma_+ + \gamma_- \sigma_-.
  \label{eq:D_general_single}
\end{align}
To simplify \eqref{eq:D_general_single}, we expand
\begin{align}
  \gamma^\dag \O \gamma
  = \abs{\gamma_\z}^2 \sigma_\z \O \sigma_\z
  + \sum_\mu \p{\abs{\gamma_\mu}^2 \sigma_\bmu \O \sigma_\mu
    + \gamma_\bmu^* \gamma_\mu \sigma_\mu \O \sigma_\mu
    + \gamma_\z^* \gamma_\mu \sigma_\z \O \sigma_\mu
    + \gamma_\bmu^* \gamma_\z \sigma_\mu \O \sigma_\z},
\end{align}
and
\begin{align}
  \gamma^\dag \gamma
  = \abs{\gamma_\z}^2
  + \sum_\mu \sp{\abs{\gamma_\mu}^2 \f12 \p{1-\mu\sigma_\z}
    + \mu \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z}
    \sigma_\mu},
\end{align}
which implies
\begin{align}
  \D\p{\gamma;\O}
  &= \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2 \D\p{\sigma_X;\O}
  + \sum_{\mu,j}
  \p{\gamma_\bmu^* \gamma_\mu \sigma_\mu^{(j)} \O \sigma_\mu^{(j)}
    + \gamma_\z^* \gamma_\mu \sigma_\z^{(j)} \O \sigma_\mu^{(j)}
    + \gamma_\bmu^* \gamma_\z \sigma_\mu^{(j)} \O \sigma_\z^{(j)}}
  \notag \\
  &\qquad -\f12 \sum_\mu \mu
  \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z} \sp{S_\mu, \O}_+.
\end{align}
In order to compute the effect of this decoherence on ordered products
of collective spin operators, we expand the anti-commutator
\begin{align}
  \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
  = S_\mu^{\ell+1} \sp{S_\z^m+\p{\mu+S_\z}^m} S_\bmu^n
  - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1}.
\end{align}
Defining
\begin{align}
  \gamma_{\z,\mu}^{(\pm)}
  &\equiv \f12\p{\gamma_\z^* \gamma_\mu \pm \gamma_\bmu^* \gamma_\z},
  \allowdisplaybreaks \\
  K_{\ell mn;\mu}
  &\equiv S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^n,
  \allowdisplaybreaks \\
  L_{\ell mn;\mu}^{(\gamma)}
  &\equiv \gamma_{\z,\mu}^{(+)} S_\mu^{\ell+1}
  \sp{\p{\mu+S_\z}^m-S_\z^m} S_\bmu^n
  + 2n \gamma_\bmu^* \gamma_z K_{\ell m,n-1;\mu}
  - 2\ell n\p{n-1} \gamma_z^* \gamma_\mu K_{\ell-1,m,n-2;\mu},
  \allowdisplaybreaks \\
  M_{\ell mn;\mu}^{(\gamma)}
  &\equiv \gamma_{\z,\mu}^{(-)} S_\mu^\ell
  S_\z^{m+1} S_\bmu^{n-1}
  -\mu \sp{\p{S-\ell-\f12n+\f12} \gamma_\z^* \gamma_\mu
    - \f12 \p{n-1} \gamma_{\z,\mu}^{(-)}}
  S_\mu^\ell S_\z^m S_\bmu^{n-1},
  \allowdisplaybreaks \\
  P_{\ell mn;\mu}
  &\equiv S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1}
  - \p{n-1} K_{\ell m,n-2;\mu},
  \allowdisplaybreaks \\
  Q_{\ell mn;\mu}^{(\gamma)}
  &\equiv \mu L_{\ell mn;\mu}^{(\gamma)}
  + 2 n M_{\ell mn;\mu}^{(\gamma)}
  + n \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu},
\end{align}
we then have
\begin{align}
  \D\p{\gamma; S_\mu^\ell S_\z^m S_\bmu^n}
  = \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2
  \D\p{\sigma_X; S_\mu^\ell S_\z^m S_\bmu^n}
  + Q_{\ell mn;\mu}^{(\gamma)} + \sp{Q_{nm\ell;\mu}^{(\gamma)}}^\dag.
\end{align}

>>>>

The most general type of single-spin decoherence is described by
  \D\p{\gamma;\O}
  = \sum_j\p{{\gamma^{(j)}}^\dag \O \gamma^{(j)}
    - \f12\sp{{\gamma^{(j)}}^\dag \gamma^{(j)}, \O}_+},
  \gamma \equiv \gamma_\z \sigma_\z
  + \gamma_+ \sigma_+ + \gamma_- \sigma_-.
  \label{eq:D_general_single}
To simplify \eqref{eq:D_general_single}, we expand
  \gamma^\dag \O \gamma
  = \abs{\gamma_\z}^2 \sigma_\z \O \sigma_\z
  + \sum_\mu \p{\abs{\gamma_\mu}^2 \sigma_\bmu \O \sigma_\mu
    + \gamma_\bmu^* \gamma_\mu \sigma_\mu \O \sigma_\mu
    + \gamma_\z^* \gamma_\mu \sigma_\z \O \sigma_\mu
    + \gamma_\bmu^* \gamma_\z \sigma_\mu \O \sigma_\z},
  \gamma^\dag \gamma
  = \abs{\gamma_\z}^2
  + \sum_\mu \sp{\abs{\gamma_\mu}^2 \f12 \p{1-\mu\sigma_\z}
    + \mu \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z}
    \sigma_\mu},
  \D\p{\gamma;\O}
  &= \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2 \D\p{\sigma_X;\O}
  + \sum_{\mu,j}
  \p{\gamma_\bmu^* \gamma_\mu \sigma_\mu^{(j)} \O \sigma_\mu^{(j)}
    + \gamma_\z^* \gamma_\mu \sigma_\z^{(j)} \O \sigma_\mu^{(j)}
    + \gamma_\bmu^* \gamma_\z \sigma_\mu^{(j)} \O \sigma_\z^{(j)}}
  &\qquad -\f12 \sum_\mu \mu
  \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z} \sp{S_\mu, \O}_+.
of collective spin operators, we expand the anti-commutator
  \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
  = S_\mu^{\ell+1} \sp{S_\z^m+\p{\mu+S_\z}^m} S_\bmu^n
  - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1}.
  \label{eq:S_mu_acomm}
Recognizing a resemblance between \eqref{eq:S_mu_acomm} and
\eqref{eq:san_z_mu}, we can collect terms to simplify
  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
  - \f{\mu}{2} \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
  = K_{\ell mn;\mu} + L_{\ell mn;\mu}
  \label{eq:dec_z_mu}
and likewise
  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
  + \f{\mu}{2} \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
  = K_{\ell mn;\mu} + M_{\ell mn;\mu}
  \label{eq:dec_mu_z}
  K_{\ell mn;\mu}
  &\equiv \f{\mu}{2} S_\mu^{\ell+1}
  \sp{\p{\mu+S_\z}^m-S_\z^m} S_\bmu^n, \\
  L_{\ell mn;\mu}
  &\equiv \mu n S_\mu^\ell \sp{-2S+2\ell+\f32\p{n-1}+\mu S_\z}
  S_\z^m S_\bmu^{n-1} \notag \\
  &\qquad - 2\mu\ell n\p{n-1} S_\mu^{\ell-1}
  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2}, \\
  M_{\ell mn;\mu}
  &\equiv \mu n S_\mu^\ell \sp{2\p{S+\mu S_\z}\p{-\mu+S_\z}^m
    - \p{\f12\sp{n-1}+\mu S_\z} S_\z^m} S_\bmu^{n-1}.
Defining for completion
  P_{\ell mn;\mu}
  \equiv \sum_j
  \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
  = n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1}
  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
  Q_{\ell mn;\mu}^{(\gamma)}
  \equiv \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu}
  + \p{\gamma_\z^* \gamma_\mu + \gamma_\bmu^* \gamma_\z}
  K_{\ell mn;\mu}
  + \gamma_\z^* \gamma_\mu L_{\ell mn;\mu}
  + \gamma_\bmu^* \gamma_\z M_{\ell mn;\mu},
  \label{eq:Q_single}
we finally have
  \D\p{\gamma; S_\mu^\ell S_\z^m S_\bmu^n}
  = \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2
  \D\p{\sigma_X; S_\mu^\ell S_\z^m S_\bmu^n}
  + Q_{\ell mn;\mu}^{(\gamma)} + \sp{Q_{nm\ell;\mu}^{(\gamma)}}^\dag.
Alternately, we can define
  \gamma_{\z,\mu}^{(\pm)}
  &\equiv \f12\p{\gamma_\z^* \gamma_\mu \pm \gamma_\bmu^* \gamma_\z},
  R_{\ell mn;\mu}
  &\equiv S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^n,
  L_{\ell mn;\mu}^{(\gamma)}
  &\equiv \mu \gamma_{\z,\mu}^{(+)} S_\mu^{\ell+1}
  \sp{\p{\mu+S_\z}^m-S_\z^m} S_\bmu^n
  + 2\mu n \gamma_\bmu^* \gamma_z R_{\ell m,n-1;\mu} \notag \\
  &\qquad - 2\mu \ell n\p{n-1} \gamma_z^* \gamma_\mu
  R_{\ell-1,m,n-2;\mu},
  M_{\ell mn;\mu}^{(\gamma)}
  &\equiv 2n \gamma_{\z,\mu}^{(-)} S_\mu^\ell S_\z^{m+1} S_\bmu^{n-1}
  &\qquad - 2\mu n \sp{\p{S-\ell-\f12n+\f12} \gamma_\z^* \gamma_\mu
    - \f12 \p{n-1} \gamma_{\z,\mu}^{(-)}}
  S_\mu^\ell S_\z^m S_\bmu^{n-1},
in terms of which
  Q_{\ell mn;\mu}^{(\gamma)}
  &\equiv L_{\ell mn;\mu}^{(\gamma)} + M_{\ell mn;\mu}^{(\gamma)}
  + \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu}.
